\section*{Abstract}

Background: Stroke is a common cause of adult disability. Expert opinion is that about 20\% of patients should receive thrombolysis to break up a clot causing the stroke. Currently, 11–12\% of patients in England and Wales receive this treatment, ranging between 2\% and 24\% between hospitals.

Objectives: By combining clinical pathway simulation and machine learning of thrombolysis decision-making and outcomes, we sought to better understand the source of variation in thrombolysis use and the effect on outcomes.

Methods: Monte-Carlo simulation was used to model the flow of patients through the emergency stroke pathway at each hospital, taking into account both average speeds and variation. XGBoost machine learning models were used to learn which patients would receive thrombolysis at each stroke team, and learn and predict outcomes with and without thrombolysis. The models were used to predict the effect of alternative improvement strategies.

Results: Combining potential changes (improving arrival-to-thrombolysis times to 30 mins, reducing ambulance call-to-hospital-arrival times by 15 minutes, having all teams attain the current upper-quartile performance in determining stroke onset time, and applying decision making similar to high thrombolysing units) would be expected to increase thrombolysis use in patients arriving by ambulance from 13\% to 20\% and double the clinical benefit (additional patients discharged mRS 0-1) from thrombolysis. The largest single factor is differences in clinical decision-making, especially around mild stroke (which make up half of all stroke admissions). Using our outcome machine learning we found that high thrombolysing are predicted to be generating more net benefit (including better avoidance of mortality and severe disability) than low thrombolylsing units. A health economics model, using the output from the outcome prediction model, estimated that thrombolysis produces an additional 0.26 QALY for each person treated.

Conclusions: There is significant inter-hospital variation in use of thrombolysis that is caused by hospital processes and differences in decision-making. Improving processes and adopting the decision making of higher-thrombolysing units would be expected to improve both thrombolysis use and outcomes. Pathway analysis and machine learning should allow stroke teams to better understand, and improve, their own pathways.